---
title: "Linear Regression in R"
---

### Plotting Linear Regression with plot()

@fig-linear_regression_inbuilt illustrates the scatterplot of the two variables (Weight vs Height) along with a linear regression line and includes summary statistics for the linear regression model including:

* **Coefficients**: Explain the meaning of the slope and intercept in the context of your data.
* **Significance Levels**: Discuss the p-values and what they indicate about the relationships.
* **R-squared value**: Explain how much of the variance in the dependent variable is explained by the model.

```{r}
#| label: fig-linear_regression_inbuilt
#| fig-cap: Weight vs Height Linear Regression Model (using plot)
#| fig-width: 8
#| fig-cap-location: top
#| fig-align: center

height <- c(151, 174, 138, 186, 128, 136, 179, 163, 152, 131)  #weight
weight <- c(63, 81, 56, 91, 47, 57, 76, 72, 62, 48) #height

# Apply the lm() function.
relation <- lm(weight~height)

plot(height,weight,col = "blue",
     abline(lm(weight~height)),cex = 1.3,pch = 16,
     xlab = "Height(cm)", ylab = "Weight(Kg)")

# print statistics
print(summary(relation))

```

### Plotting Linear Regression with plot()

@fig-linear_regression_ggplot illustrates the same data as @fig-linear_regression_inbuilt but uses ggplot() to plot the output of the model.

```{r}
#| label: fig-linear_regression_ggplot
#| fig-cap: Weight vs Height Linear Regression Model (using ggplot)
#| fig-width: 8
#| fig-cap-location: top
#| fig-align: center

# convert to a dataframe to use ggplot()
data <- data.frame(
  height <- c(151, 174, 138, 186, 128, 136, 179, 163, 152, 131),
  weight <- c(63, 81, 56, 91, 47, 57, 76, 72, 62, 48)
)

# Run linear regression
model <- lm(weight~height, data = data)

# View the summary of the model
summary(model)

#using ggplot
p <- ggplot(data, aes(x = height, y = weight)) +
  geom_point() +
  labs(
    x = "Height(cm)",
    y = "Weight(Kg)"
  ) +
  theme_minimal()

# Add the linear regression line
p + geom_smooth(method = "lm", se = FALSE, color = "blue")

```




